---
title: "Data Cleaning"
author: "ARC Capstone Team"
date: "4/5/2022"
output: html_document
---

## Import and Load Package
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = setdiff(ls(), lsf.str()))
install.packages("tidyverse")
library(tidyverse)
library(dplyr)
library(tidyverse)
```

# Setting up directory
```{r}

# Name of the folder where csv files are located
csv_dir <- "rawdata"

# Location and file name of the csv file 
eviction_filing_csv <- "evictions.csv"
household_count_csv <- "household_count.csv"
education_19_csv <- "education_2019.csv"
education_20_csv <- "education_2020.csv"
rent_burden_19_csv <- "rent_burden_2019.csv"
rent_burden_20_csv <- "rent_burden_2020.csv"
unemployment_19_csv <- "unemp_2019.csv"
unemployment_20_csv <- "unemp_2020.csv"
uninsured_19_csv <- "uninsured_2019.csv"
uninsured_20_csv <- "uninsured_2020.csv"
race_18_19_csv <- "race_18_19.csv"
race_20_csv <- "race_2020.csv"
poverty_18_19_csv <- "poverty_18_19.csv"
poverty_20_csv <- "poverty_2020.csv"
renter_occupied_19_csv <- "renter_occupied_2019.csv"
renter_occupied_20_csv <- "renter_occupied_2020.csv"

# Location and file name of the csv file 
eviction_filing_dir <- file.path(csv_dir, eviction_filing_csv)
household_count_dir <- file.path(csv_dir, household_count_csv)
education_19_dir <- file.path(csv_dir, education_19_csv)
education_20_dir <- file.path(csv_dir, education_20_csv)
rent_burden_19_dir <- file.path(csv_dir, rent_burden_19_csv)
rent_burden_20_dir <- file.path(csv_dir, rent_burden_20_csv)
unemployment_19_dir <- file.path(csv_dir, unemployment_19_csv)
unemployment_20_dir <- file.path(csv_dir, unemployment_20_csv)
uninsured_19_dir <- file.path(csv_dir, uninsured_19_csv)
uninsured_20_dir <- file.path(csv_dir, uninsured_20_csv)
race_18_19_dir <- file.path(csv_dir, race_18_19_csv)
race_20_dir <- file.path(csv_dir, race_20_csv)
poverty_18_19_dir <- file.path(csv_dir, poverty_18_19_csv)
poverty_20_dir <-  file.path(csv_dir, poverty_20_csv)
renter_occupied_19_dir <- file.path(csv_dir, renter_occupied_19_csv)
renter_occupied_20_dir <- file.path(csv_dir, renter_occupied_20_csv)

# A folder where cleaned dataset will be saved
cleaned_csv_folder <- "cleaned_data"
dir.create(cleaned_csv_folder, showWarnings = FALSE)

# A folder where cleaned dataset will be saved
final_csv_folder <- "final_csv"
dir.create(final_csv_folder, showWarnings = FALSE)

```

```

## Import Raw Datasets
```{r}
# Import poverty dataset of year 2020 from the U.S. Census Bureau website
pov2020 <- read.csv(poverty_20_dir)
# Import poverty dataset of year 2018 and 2019 from the Neighborhood Nexus website
pov1819 <- read.csv(poverty_18_19_dir)

# Import education data  that estimates education attainment
edu_20 <-read.csv(education_19_dir)
edu_19 <- read.csv(education_20_dir)

# Import 2019 and 2020 rent burden data 
rent_burden_19 <- read.csv(rent_burden_19_dir)
rent_burden_20 <- read.csv(rent_burden_20_dir)

# Import uninsurance rate datasets of years 2019 and 2020 from the U.S. Census Bureau website
uninsurance19<-read.csv(uninsured_19_dir)
uninsurance20<-read.csv(uninsured_20_dir)

# Import race dataset of year 2020 from the U.S. Census Bureau website
race2020 <- read.csv(race_20_dir)
# Import race dataset of year 2018 and 2019 from the Neighborhood Nexus website
race1819 <- read.csv(race_18_19_dir)

## Import renter occupied housing unit data from the U.S. Census Bureau website for 2019
rent19=read.csv(renter_occupied_19_dir)

#Import unemployment rate datasets of years 2019 and 2020 from the U.S. Census Bureau website
unemp19<-read.csv(unemployment_19_dir)
unemp20<-read.csv(unemployment_20_dir)
```



## Extract and Clean X Variables
```{r}
### Poverty 
### 2020

# Select GEO_ID, NAME, percent in poverty columns for 2020 data
pov_2020 <- pov2020[-1,] %>%
  rename("TractID" = "GEO_ID", "PovertyPercent" = "S1701_C03_001E") %>%
  mutate("PovertyRate" = as.numeric(as.character(PovertyPercent))/100) %>% mutate(TractID = substr(TractID, 10, 20)) %>%
  select("TractID", "PovertyRate") 

# Export 2020 race data as a csv file
write.csv(pov_2020, file = file.path(cleaned_csv_folder, "poverty_20_cleaned.csv"), row.names=FALSE)

### 2019 

pov19 <- pov1819[,c(1,3,6)]
names(pov19) <- c("TractID", "COUNTY", "PovertyPercent")
# Filter 5 counties and rename columns
pov19 <- pov19 %>%
  filter(COUNTY %in% c("Fulton", "DeKalb", "Gwinnett", "Clayton", "Cobb")) %>% 
  mutate("PovertyRate" = as.numeric(as.character(PovertyPercent))/100) %>% select(TractID, COUNTY, PovertyRate)
 
# Export 2018, 2019 poverty data as a csv file
write.csv(pov19, file.path(cleaned_csv_folder, "poverty_19_cleaned.csv"), row.names=FALSE)


### Education - Rate of people who got bach degree or higher
### 2019 and 2020

# Define a function which cleans the educaton data
education_cleaning <- function(df)  { 
edu <- df %>%
  select(GEO_ID, NAME, S2301_C01_035E, S2301_C01_021E)
# Drop the first row which contains description of the data
# and save it as a new dataframe
edu_new <- edu[-1,]
# Create a column that contains # of people who got bach degree
edu_new$bach<-as.numeric(edu_new$S2301_C01_035E)
# Create a column that contains population living in different census tracts
edu_new$pop<-as.numeric(edu_new$S2301_C01_021E)
# Create a column that contains percentage of ppl who got bach degrees or higher
edu_new$education_rate_bachelors_or_higher <- edu_new$bach/edu_new$pop
# Create a dataframe that only contains columns that are relevant to our analyses
edu_new1 <- edu_new %>% mutate(GEO_ID = substr(GEO_ID, 10, 20)) %>%
  select(GEO_ID, education_rate_bachelors_or_higher) 
# Rename columns
colnames(edu_new1) <- c('TractID', 'EducationRate')
return(edu_new1)
}
# Export 2020 education data as a csv file
education_rate_20 <- education_cleaning(edu_20)
write.csv(education_rate_20, file.path(cleaned_csv_folder, "education_rate_cleaned_2020.csv"), row.names=FALSE)
# Export 2019 education data as a csv file
education_rate_19 <- education_cleaning(edu_19)
write.csv(education_rate_19, file.path(cleaned_csv_folder, "education_rate_cleaned_2019.csv"), row.names=FALSE)


### Rent Burden
### 2019

# Rename variable
names(rent_burden_19)[6] <- "RentBurdenPerc"
# Filter 5 county area
rent_burden_19 <- rent_burden_19 %>%
  filter(COUNTY == "Fulton" | COUNTY == "DeKalb"| COUNTY == "Cobb" | COUNTY == "Clayton" | COUNTY == "Gwinnett") %>% mutate(RentBurdenRate = RentBurdenPerc/100)
  
# rename columns
rent_burden_19 <- rename(rent_burden_19,TractID=GEOID)
# Drop columns that are irrelevant to analyses
rent_burden_19 <- rent_burden_19 %>% select(TractID, COUNTY, RentBurdenRate)
# Export cleaned rent burden data for 2019
write.csv(rent_burden_19, file.path(cleaned_csv_folder, "rent_burden_rate_cleaned_2019.csv"), row.names=FALSE)

### 2020

# Select GEO_ID, NAME, and columns that shows rent payment consist 30 percent or more of household income
new_20 <- rent_burden_20 %>%
  select(B25070_001E, B25070_007E, B25070_008E, B25070_009E, B25070_010E, GEO_ID, NAME)

# Rename columns
new_20_1 <- new_20[-1,]
new_20_1$total<-as.numeric(new_20_1$B25070_001E)
new_20_1$rent_30<-as.numeric(new_20_1$B25070_007E)
new_20_1$rent_35<-as.numeric(new_20_1$B25070_008E)
new_20_1$rent_40<-as.numeric(new_20_1$B25070_009E)
new_20_1$rent_50<-as.numeric(new_20_1$B25070_010E)

# Calculate the rent burden percentage
new_20_1$sum <- new_20_1$rent_30+new_20_1$rent_35+new_20_1$rent_40+new_20_1$rent_50
new_20_1$percentage <- new_20_1$sum/new_20_1$total

rent_burden_20 <- new_20_1 %>% mutate(TractID = substr(GEO_ID, 10, 20))  %>%
  select(TractID, percentage) 
rent_burden_20 <- rename(rent_burden_20,RentBurdenRate=percentage)

# Export cleaned rent burden data for 2020
write.csv(rent_burden_20, file.path(cleaned_csv_folder, "rent_burden_rate_cleaned_2020.csv"), row.names=FALSE)

### Uninsured rate
### 2019 & 2020

# Define a function that drops, renames column
uninsur_cleaning <- function(uninsur){
  uninsur=uninsur[-c(1),]
  uninsur=uninsur %>% select(GEO_ID, NAME, S2701_C05_001E)
  uninsur=uninsur%>%mutate(TractID=substr(GEO_ID,10,20), UninsurRate = as.numeric(as.character(S2701_C05_001E))/100)
  uninsur=uninsur%>%select(TractID,UninsurRate)
  return(uninsur)
}

uninsurance_19=uninsur_cleaning(uninsurance19)
uninsurance_20=uninsur_cleaning(uninsurance20)

# Export 2019 unemployment data as a csv file
write.csv(uninsurance_19, file.path(cleaned_csv_folder ,"uninsurance_rate_cleaned_2019.csv"), row.names=FALSE)
# Export 2020 unemployment data as a csv file
write.csv(uninsurance_20, file.path(cleaned_csv_folder ,"uninsurance_rate_cleaned_2020.csv"), row.names=FALSE)


### Minority: Percentage of non-white population
### 2020

# Select GEO_ID, NAME, total population, and the white alone population columns for 2019 data
# Calculate Minority percentage
race_2020 <- race2020[-c(1),] %>%
  select("GEO_ID", "NAME", "P1_001N", "P1_003N") %>%
  mutate(MinorityRate = (as.numeric(as.character(P1_001N)) - as.numeric(as.character(P1_003N)))/ as.numeric(as.character(P1_001N))) %>% mutate(TractID=substr(GEO_ID,10,20)) %>% select(TractID, MinorityRate)

# Export 2020 race data as a csv file
write.csv(race_2020, file = file.path(cleaned_csv_folder, "race_20_cleaned.csv"), row.names=FALSE)

### 2019
race19 <- race1819[,c(1,3,6)]
names(race19) <- c("TractID", "COUNTY", "MinorityPercent")
race_19 <- race19 %>%
  filter(COUNTY %in% c("Fulton", "DeKalb", "Gwinnett", "Clayton", "Cobb")) %>% 
  mutate("MinorityRate" = as.numeric(as.character(MinorityPercent))/100) %>%
  select("TractID", "COUNTY", "MinorityRate")

# Export 18-19 race data as a csv file
write.csv(race_19, file =  file.path(cleaned_csv_folder, "race_19_cleaned.csv"), row.names=FALSE)

### Renter Occupied
### 2019 

## Remove the first row which contains the header
rent19=rent19[-c(1),]
## Calculate the renter occupied housing rate and rename it as renter_rate for 2019 data
rent19=rent19%>%mutate(RenterRate=as.numeric(B25003_003E)/as.numeric(B25003_001E))
rent19=rent19%>%select(GEO_ID,RenterRate)
rent19=rent19%>%mutate(TractID=substr(GEO_ID,10,20))
rent19=rent19%>%select(TractID,RenterRate)

## Import renter occupied housing unit data from the U.S. Census Bureau website for 2020
rent20=read.csv(renter_occupied_20_dir)
rent20=rent20[-c(1),]

## Calculate the renter occupied housing rate and rename it as renter_rate for 2020 data
rent20=rent20%>%mutate(RenterRate=as.numeric(B25003_003E)/as.numeric(B25003_001E))
rent20=rent20%>%select(GEO_ID,RenterRate)
rent20=rent20%>%mutate(TractID=substr(GEO_ID,10,20))
rent20=rent20%>%select(TractID,RenterRate)

## Export 2020 renter occupied housing rate
write.csv(rent19, file.path(cleaned_csv_folder, "renter_occupied_cleaned_2019.csv"), row.names=FALSE)
write.csv(rent20, file.path(cleaned_csv_folder, "renter_occupied_cleaned_2020.csv"), row.names=FALSE)

### Unemployment Rate
### 2019 & 2020

# Define a function that cleans unemployment csv files
unemp_cleaning <- function(unemp){
  unemp <- unemp%>%select(GEO_ID, NAME, S2301_C04_001E)
  unemp <- unemp[-1, ] %>% mutate("TractID"=substr(GEO_ID,10,20), 
                                  "UnempRate"=as.numeric(as.character(S2301_C04_001E))/100)
  unemp <- unemp %>% select(TractID, UnempRate)
  return(unemp)
}

# Export 2019 unemployment data as a csv file
write.csv(unemp_cleaning(unemp19), file.path(cleaned_csv_folder, "unemployment_rate_cleaned_2019.csv"), row.names=FALSE)
# Export 2020 unemployment data as a csv file
write.csv(unemp_cleaning(unemp20), file.path(cleaned_csv_folder, "unemployment_rate_cleaned_2020.csv"), row.names=FALSE)
```


## Extract and Clean Y Variables 
# Compute annual/monthly eviction rate
```{r}
# import household count data from neighborhood nexus
household <- read.csv(household_count_dir, skip = 1)
# import eviction filing data
eviction  <- read.csv(eviction_filing_dir)

compute_evic_rate <- function(eviction_df, household_df, annual = TRUE){
  # Split filing date into days, month, and years
  eviction_df <- separate(data = eviction, col = FilingDate, sep = "/", 
                 into = c("Dates","Month","Year"))
  if (annual){# Compute annual total eviction filings
      df <- eviction_df%>% group_by(TractID, Year) %>% 
  summarize("TotalFilings" = sum(TotalFilings), )
    
  }
    else {# If annual == FALSE, compute monthly total eviction filings
      df <- df %>% group_by(TractID, CountyID, Year, Month) %>% 
  summarize("TotalFilings" = sum(TotalFilings),)}
  
  # Remove observations that are missing census tract ID
  df <- df[df$TractID != 99999999999, ]
  # Rename variable of household data 
  names(household_df)[6] <- "TotalHousehold2019"
  # Subset household_df
  df2 <- household_df[, c("GEOID", "TotalHousehold2019")]
  # Merge eviction filing data with household count data
  df3 <- inner_join(df, df2, by=c("TractID" = "GEOID"))
  # Compute eviction rate
  df3 <- df3 %>% mutate("EvictionRate" = TotalFilings/TotalHousehold2019)
  # Remove observations with NAs
  df3 <- na.omit(df3)
  return(df3)
}
```

## Merge Datasets
```{r}
# Merge 2019
## import Y dataset
y=read.csv(file.path(cleaned_csv_folder, "monthly_eviction_rate_from_2019_to_2022.csv"))
y2019=y%>%filter(Year==19)

## import poverty rate data
pov=read.csv(file.path(cleaned_csv_folder, "poverty_19_cleaned.csv"))

## import educ data 
educ=read.csv(file.path(cleaned_csv_folder, "education_rate_cleaned_2019.csv"))

## import rent burden
rent=read_csv(file.path(cleaned_csv_folder, "rent_burden_rate_cleaned_2019.csv"))

## import insurance data
uninsur=read_csv(file.path(cleaned_csv_folder, "uninsurance_rate_cleaned_2019.csv"))

## import race data
mino=read.csv(file.path(cleaned_csv_folder, "race_19_cleaned.csv"))

## import renter occupied percent
renter <- read.csv(file.path(cleaned_csv_folder, "renter_occupied_cleaned_2019.csv"))

## import Unemployment rate 
unemp <- read.csv(file.path(cleaned_csv_folder, "unemployment_rate_cleaned_2019.csv"))


tract_to_char <- function(df){
  df$TractID <- as.character(df$TractID)
  return(df)
}

y2019=tract_to_char(y2019)
pov=tract_to_char(pov)
educ=tract_to_char(educ)
rent=tract_to_char(rent)
uninsur=tract_to_char(uninsur)
mino=tract_to_char(mino)
renter=tract_to_char(renter)
tract=tract_to_char(unemp)


df_2019 <- merge(y2019, pov)
df_2019 <- merge(df_2019, educ)
df_2019 <- merge(df_2019, rent)
df_2019 <- merge(df_2019, uninsur)
df_2019 <- merge(df_2019, mino )
df_2019 <- merge(df_2019, renter)
df_2019 <- merge(df_2019, unemp)
write.csv(df_2019, file.path(final_csv_folder, "final_data_2019.csv"))
```

```{r}
# Merge 2020
## import Y dataset
y=read.csv(file.path(cleaned_csv_folder, "monthly_eviction_rate_from_2019_to_2022.csv"))
y2020=y%>%filter(Year==20)

## import poverty rate data
pov=read.csv(file.path(cleaned_csv_folder, "poverty_20_cleaned.csv"))

## import educ data 
educ=read.csv(file.path(cleaned_csv_folder, "education_rate_cleaned_2020.csv"))

## import rent burden
rent=read_csv(file.path(cleaned_csv_folder, "rent_burden_rate_cleaned_2020.csv"))

## import insurance data
uninsur=read_csv(file.path(cleaned_csv_folder, "uninsurance_rate_cleaned_2020.csv"))

## import race data
mino=read.csv(file.path(cleaned_csv_folder, "race_20_cleaned.csv"))

## import renter occupied percent
renter <- read.csv(file.path(cleaned_csv_folder, "renter_occupied_cleaned_2020.csv"))

## import Unemployment rate 
unemp <- read.csv(file.path(cleaned_csv_folder, "unemployment_rate_cleaned_2020.csv"))


tract_to_char <- function(df){
  df$TractID <- as.character(df$TractID)
  return(df)
}

y2020=tract_to_char(y2020)
pov=tract_to_char(pov)
educ=tract_to_char(educ)
rent=tract_to_char(rent)
uninsur=tract_to_char(uninsur)
mino=tract_to_char(mino)
renter=tract_to_char(renter)
tract=tract_to_char(unemp)


df_2020 <- merge(y2020, pov)
nrow(df_2020)
df_2020 <- merge(df_2020, educ)
df_2020 <- merge(df_2020, rent)
nrow(df_2020)
df_2020 <- merge(df_2020, uninsur)
df_2020 <- merge(df_2020, mino )
df_2020 <- merge(df_2020, renter)
df_2020 <- merge(df_2020, unemp)
write.csv(df_2020, file.path(final_csv_folder, "final_data_2020.csv"))
```




